---
title: "Hands-on_Exercise02"
format: html
---

# Overview

This chapter introduces ggplot2 extensions for refined statistical graphics, including annotation controlwith ggrepel, publication-quality themes via ggthemes and hrbrthemes, and composite plotting with patchwork.

# Getting Started

## Installing and loading the required libraries

Code chunk below will be used to check if these packages have been installed and also will load themonto the working R environment.

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

### **Importing data**

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## **Beyond ggplot2 Annotation: ggrepel**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### **Working with ggrepel**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

## **Beyond ggplot2 Themes**

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 
```

### **Working with ggtheme package**

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()
```

### **Working with hrbthems package**

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = "grey25",
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_minimal()
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 boundary = 100,
                 color = "grey25",
                 fill = "grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_minimal(base_size = 15) +
  theme(
    axis.title = element_text(size = 18),
    panel.grid.major.y = element_line(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank()
  )
```

## **Beyond Single Graph**

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")
```

### **Creating Composite Graphics: pathwork methods**

### **Combining two ggplot2 graphs**

```{r}
p1 + p2
```

### **Combining three ggplot2 graphs**

```{r}
(p1 / p2) | p3
```

### **Creating a composite figure with tag**

```{r}
((p1 / p2) | p3) + 
  plot_annotation(tag_levels = 'I')
```

### **Creating figure with insert**

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.7, 
                   right = 0.5, 
                   top = 1)
```

### **Creating a composite figure by using patchwork and ggtheme**

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_economist()
```
