---
title: "Hands-on Exercise01"
format: html
---

# overview

This chapter introduces the core concepts of ggplot2, emphasizing the layered grammar of graphics and the practical application of its essential components to create effective and interpretable statistical visualisations.

# Getting Started

### **Installing and loading the required libraries**

```{r}
pacman::p_load(tidyverse)
```

### **Importing data**

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

### **R Graphics VS ggplot**

```{r}
ggplot(data=exam_data, aes(x = MATHS)) +
  geom_histogram(bins=10, 
                 boundary = 100,
                 color="black", 
                 fill="grey") +
  ggtitle("Distribution of Maths scores")
```

## **Essential Grammatical Elements in ggplot2: data**

```{r}
ggplot(data=exam_data)
```

## **Essential Grammatical Elements in ggplot2:**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS))
```

### **Geometric Objects: geom_bar**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

### **Geometric Objects: geom_dotplot**

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,         
               dotsize = 0.5) +      
  scale_y_continuous(NULL,           
                     breaks = NULL)  
```

### **Geometric Objects: `geom_histogram()`**

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()       
```

### **Modifying a geometric object by changing `geom()`**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

### **Modifying a geometric object by changing *aes()***

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```

### **Geometric Objects: geom-density()**

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()
```

### **Geometric Objects: geom_boxplot**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```

### **Geometric Objects: geom_violin**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()
```

### **Geometric Objects: geom_point()**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()            
```

### ***geom*** **objects can be combined**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", 
             size = 0.5)        
```

## **Essential Grammatical Elements in ggplot2: stat**

There are two ways to use these functions:

-   add a `stat_()` function and override the default geom, or

-   add a `geom_()` function and override the default stat.

### **Working with `stat()`**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot()
```

### **Working with stat - the *stat_summary()* method**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun = "mean",         
               colour ="red",        
               size=4)               
```

### **Working with stat - the `geom()` method**

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun="mean",           
             colour="red",          
             size=4)          
```

### **Adding a best fit curve on a scatterplot?**

```{R}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(size=0.5)
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              linewidth=0.5)
```

## **Essential Grammatical Elements in ggplot2: Facets**

### **Working with `facet_wrap()`**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_wrap(~ CLASS)
```

### **`facet_grid()` function**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20) +
    facet_grid(~ CLASS)
```

## **Essential Grammatical Elements in ggplot2: Coordinates**

### **Working with Coordinate**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip()
```

### **Changing the y- and x-axis range**

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, size=0.5)
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
```

## **Essential Grammatical Elements in ggplot2: themes**

### **Working with theme**

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_gray()
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_classic()
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal()
```
